Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk

### 2021-09-27 VegOzone.tex
Remove subsections

\subsubsection{Temperature, precipitation, and global irradiance}
\label{subsubsec:clim_temp_prec}
The location of Svanvik ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate. The derived climate diagram based on climatological data (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well. Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred July 16 the same year.

@@

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscandia. To this end, we first calculate a pairwise correlation between the different stations. The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}). Therefore, we select these data to compute the generalized ozone climatology for northern Fennoscandia. The correlation between Svanvik with any of the active monitoring sites in 2018 (Esrange and Pallas) is rather low, 0.4 and 0.6, respectively. Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.

In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined data (Esrange, Jergul/Karasjok, Pallas) (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown. Darker colors indicate higher probability to observe these values. On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison.

The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in the beginning of September ($240-250\,\unit{doy}$) and amounts to $24-30\,\unit{ppb}$. The turning points lie approximately in June and November. The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum. The most pronounced increase in \chem{\left<[O_3]\right>} coincides with rising temperatures and global irradiance. The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.

The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generalized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude. The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}. The estimated average begin of growing season for deciduous trees ($\left<G_\mathrm{start}\right>_\mathrm{decid}$) is $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details). In July--September, ozone is occasionally almost completely depleted during nighttime. This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air. The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the time series taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).

@@

\subsubsection{Temperature, precipitation, and global irradiance}
\label{subsubsec:anomal_tpq}
In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month. Negative deviations from climatology are displayed as a negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). We find, that the summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer. Significantly warmer days continued to occur in all months (August--November) except for October. March 2018 had many unusually cold days. The summer of 2019 was fairly average, however, significantly colder days occurred in July, while August had both significantly colder and warmer days. April 2019 had significantly warmer than average days.

@@

\subsubsection{Ozone}
\label{subsubsec:anomal_ozone}
\textbf{TODO: Remove all content which is part of the technical note}
To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018. To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper). In the following, we will describe the bias-correction and discuss the significance of \chem{[O_3]} enhancements in 2018.

>>>>

The location of Svanvik ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate. The derived climate diagram based on climatological data (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well. Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred July 16 the same year.

@@

The derived ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}. Darker colors indicate higher probability to observe these values. On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison. The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a pronounced seasonal cycle. The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude. The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}. The estimated average begin of growing season for deciduous trees ($\left<G_\mathrm{start}\right>_\mathrm{decid}$) is $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details). In July--September, ozone is occasionally almost completely depleted during nighttime. This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air. The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the time series taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).

@@

In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month. Negative deviations from climatology are displayed as a negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). We find, that the summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer. Significantly warmer days continued to occur in all months (August--November) except for October. March 2018 had many unusually cold days. The summer of 2019 was fairly average, however, significantly colder days occurred in July, while August had both significantly colder and warmer days. April 2019 had significantly warmer than average days.

@@

\textbf{TODO: Remove all content which is part of the technical note}
To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018. To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper). In the following, we will describe the bias-correction and discuss the significance of \chem{[O_3]} enhancements in 2018.
